Interval Probability for Fuzzy Quantum Theories

### 2016-05-11 prob.tex
prob.tex \subsection{Buffon's Needle Problem} Update the sources...

Suppose we drop a needle of length $\ell$ onto a floor made of equally
spaced parallel lines a distance $\ell$ apart. It is a known fact that
the probability of the needle crossing a line converges to
$\frac{2}{\pi} = 0.6366197723675814$~\cite{Buffon1777}.
%% \yutsung{Add $\ldots$ at the end? }\\
We analyze this situation in the mathematical framework of probability
spaces paying special attention to the resources needed to estimate
the probability computationally or experimentally.

To formalize the experiment, we consider an experimental setup
consisting of a collection of $N$ identical needles of length
$\ell$. We throw the $N$ needles one needle at a time, and observe the number
$X$ of needles that cross a line. The sample space can be expressed as the set
$\{X,-\}^N$
%% \yutsung{Use more physical or mathematical notation?}\\
of sequences of characters of length $N$ where each character is
either $X$ to indicate a needle crossing a line or $-$ to indicate a
needle not crossing a line. If $N=3$, the event that exactly 2 needles
cross lines~$\{-XX,X-X,XX-\}$ has
probability~$\frac{3}{8}$. Generally, the event that exactly $M$
needles out of the $N$ total needles cross lines is $\frac{M}{N}$.

In an actual experiment with 500 needles, it was found that 321
crossed a line which gives a probability of $0.642$. In a larger
experiment with $3408$ needles, the probability was calculated to be
$0.6366197193098592$~\cite{Lazzarini1901,Uspenskyothers1937}.\\
\yutsung{
May I confirm how many times needle cross the line within $3408$ needles?
Because $\frac{2169}{3408}$ is almost $0.6364436619718310$, and
$\frac{2170}{3408}$ is almost $0.6367370892018779$, both of them are not
close to $0.6366197193098592$ enough\ldots 
}\\
\amr{
There are several articles and papers on the experiment. I got these
values from wikipedia and mathworld. We need to get precise references
and cite the papers describing the experiments. The actual values
don't matter though.
}\\
\yutsung{
Although I don't know Italian, it seems that Lazzarini~\cite{Lazzarini1901}
used the needle with length 2.5, and the length among lines is 3.
When he tossed needles 3408 times, he got 1808 crossing, and 
\[
\pi\approx\frac{2\cdot2.5\cdot3408}{3\cdot1808}=\frac{355}{113}\approx3.1415929
\]
However, his paper was be criticized in Wikipedia, because $\frac{355}{113}$
is the best rational approximation of $\pi$ for small intergers.
How can his experiment just gave this number?

Uspensky~\cite{Uspenskyothers1937} and Buffon~\cite{Buffon1777}
are books. I will check out \cite{Uspenskyothers1937} on Sunday,
and \cite{Buffon1777} on Monday...
}\\
Comparing the values to the idealized mathematical value, we see that
the observed probability approaches $\frac{2}{\pi}$ but only if
\emph{larger and larger resources} are expended.

These resource considerations thus lead us to replace the real
interval $[0,1]$ with rational numbers up to a certain precision.  The
intuition is that the precision depends on the amount of resources
allocated for the measurement process. There is clearly a connection
between the number of needles and the achievable precision: in the
hypothetical experiment with 3 needles, it is not sensible to retain
100 digits in the expansion of $\frac{2}{\pi}$.

There is however another more subtle source of unbounded computational
power in the experiment. We are assuming that we can always determine
with certainty whether a needle is crossing a line. But ``lines'' on
the the floor have thickness, their distance apart is not exactly
$\ell$. and the needles lengths are not all equal and are not all
absolutely equal to $\ell$. These perturbations make the events
``fuzzy.'' For example, instead of talking about the idealized event
that exactly $M$ needles cross lines which would require the most
expensive needles built to the most precise accuracy, laser precision
for drawing lines on the floor, and the most powerful microscopes to
determine if a needle does cross a line, we might instead talk about
the event that $M-\delta$ needles evidently cross lines and
$M+\delta'$ needles plausibly cross lines for some acceptable
approximations $\delta$ and $\delta'$. This fuzzy notion of events
leads us to consider intervals of probability reflecting the certainty
of the event and its plausibility. In the hypothetical experiment with
3 needles, assume that one needle clearly crossed a line and another
was close enough to plausibly be considered as crossing the line: we
would express the probability in this case as the interval
$[\frac{1}{3},\frac{2}{3}]$ expressing that we are certain that the
event has probability at least $\frac{1}{3}$ but it is possible that
it would have probability $\frac{2}{3}$.

>>>>

Suppose we drop a needle of length $\ell$ onto a floor made of equally
spaced parallel lines a distance $h$ apart. It is a known fact that
the probability of the needle crossing a line converges to
$\frac{2\ell}{\pi h}$~\cite{Buffon1777,DeMorgan1872,Hall1873,Uspensky1937}.
We analyze this situation in the mathematical framework of probability
spaces paying special attention to the resources needed to estimate
the probability computationally or experimentally.

To formalize the experiment, we consider an experimental setup
consisting of a collection of $N$ identical needles of length
$\ell$. We throw the $N$ needles one needle at a time, and observe the number
$X$ of needles that cross a line. The sample space can be expressed as the set
$\{X,-\}^N$
of sequences of characters of length $N$ where each character is
either $X$ to indicate a needle crossing a line or $-$ to indicate a
needle not crossing a line. If $N=3$, the probability of the event that exactly
2 needles cross lines~$\{-XX,X-X,XX-\}$ can be 
estimated by the relative frequency~$\frac{3}{8}$. Generally, the probability of
the event that exactly $M$ needles out of the $N$ total needles cross lines can
be estimated by the relative frequency~$\frac{M}{N}$.

In an actual experiment with $500$ needles and the ratio
$\frac{\ell}{h}=0.75$~\cite{Hall1873}, it was found that $236$ 
crossed a line so the relative frequency is $0.472$ comparing 
the values to the idealized mathematical probability $0.4775\ldots$. In a larger 
experiment with $5000$ needles and the ratio $\frac{\ell}{h}=0.8$~\cite{Uspensky1937}, 
the relative frequency was calculated to be $0.5064$ comparing 
the values to the idealized mathematical probability $0.5093\ldots$. We see that 
the observed probability approaches $\frac{2\ell}{\pi h}$ but only if
\emph{larger and larger resources} are expended.

These resource considerations thus lead us to replace the real
interval $[0,1]$ with rational numbers up to a certain precision.  The
intuition is that the precision depends on the amount of resources
allocated for the measurement process. There is clearly a connection
between the number of needles and the achievable precision: in the
hypothetical experiment with 3 needles, it is not sensible to retain
100 digits in the expansion of $\frac{2\ell}{\pi h}$.

There is however another more subtle source of unbounded computational
power in the experiment. We are assuming that we can always determine
with certainty whether a needle is crossing a line. But ``lines'' on
the the floor have thickness, their distance apart is not exactly
$\ell$. and the needles lengths are not all equal and are not all
absolutely equal to $\ell$. These perturbations make the events
``fuzzy.'' For example, instead of talking about the idealized event
that exactly $M$ needles cross lines which would require the most
expensive needles built to the most precise accuracy, laser precision
for drawing lines on the floor, and the most powerful microscopes to
determine if a needle does cross a line, we might instead talk about
the event that $M-\delta$ needles evidently cross lines and
$M+\delta'$ needles plausibly cross lines for some acceptable
approximations $\delta$ and $\delta'$. This fuzzy notion of events
leads us to consider intervals of probability reflecting the certainty
of the event and its plausibility. In an actual experiment with
$3204$ needles and the ratio $\frac{\ell}{h}=0.6$~\cite{DeMorgan1872}, 
there are $1213$ needles clearly crossed a line and $11$ ones
was close enough to plausibly be considered as crossing the line: we
would express the probability in this case as the interval
$\left[\frac{1213}{3204},\frac{1234}{3204}\right]$ expressing that we are
certain that the event has probability at least $\frac{1213}{3204}$ but it is possible that
it would have probability $\frac{1234}{3204}$.
